---
title: "Assignment7"
author: "Chhiring Lama"
date: "2024-03-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Books Info in 3 different file formats

Load libraries

```{r , echo=TRUE}
library(dplyr)
library(tidyr)
library(xml2)
library(rjson)
library(XML)
library(DT)
library(rvest)
library(tibble)
library(jsonlite)

```

# Load html file 

```{r, echo=TRUE}
book_info_html <- read_html("https://raw.githubusercontent.com/topkelama/lfsStorage/main/info_books.html")
```

Convert to dataframe

```{r, echo=TRUE}
tables1 <- html_table(book_info_html)
book_df_html <- data.frame(tables1[[1]], stringsAsFactors = FALSE)
book_df_html
```
# Load xml file

```{r, echo=TRUE}
book_info_xml <- read_xml("https://raw.githubusercontent.com/topkelama/lfsStorage/main/info_books.xml")
```

convert to dataframe

```{r, echo=TRUE}
bookInfo_parse <- xmlParse(book_info_xml)
book_xml_df <- xmlToDataFrame(bookInfo_parse, stringsAsFactors = FALSE)

```

```{r, echo=TRUE}
book_xml_df
```


```{r, echo=FALSE}
# # Extract authors from XML data
# Authors <- lapply(xml_find_all(book_info_xml, xpath = "//Authors"), function(authors) {
#   xml_children(authors) %>% 
#     xml_text() %>% 
#     trimws()
# })
# 
# Pages <- as.integer(xml_text(xml_find_all(book_info_xml, xpath = "//Pages")))
# Release_Year <- as.integer(xml_text(xml_find_all(book_info_xml, xpath = "//Release_Year")))
# Price <- as.numeric(xml_text(xml_find_all(book_info_xml, xpath = "//Price")))
# ISBN <- xml_text(xml_find_all(book_info_xml, xpath = "//ISBN"))
```


```{r, echo=FALSE}
# # Create data frame
# book_df <- data.frame(
#   Title = rep(Title, lengths(Authors)),
#   Authors = unlist(Authors),
#   Pages = rep(Pages, lengths(Authors)),
#   Release_Year = rep(Release_Year, lengths(Authors)),
#   Price = rep(Price, lengths(Authors)),
#   ISBN = rep(ISBN, lengths(Authors))
# )
# 
# # Print the data frame
# print(book_df)
```

# Load json file

```{r, echo=TRUE}
json_file <- "https://raw.githubusercontent.com/topkelama/lfsStorage/main/info_books.json"
download.file(json_file, destfile = "info_books.json", quiet = TRUE)
```


```{r, echo=TRUE}
# Read and parse the JSON file
book_info_json <- fromJSON("info_books.json")
```


```{r, echo=TRUE}
# Expand the Authors list into separate rows
book_info_json <- book_info_json %>%
  mutate(Authors = lapply(Authors, toString)) %>%
  unnest(Authors)
book_info_json
```



```{r, echo=FALSE}
# # Create a data frame
# book_df_json <- data.frame(
#   Title = rep(book_info_json$Title, sapply(book_info_json$Authors, length)),
#   Authors = unlist(book_info_json$Authors),
#   Pages = rep(book_info_json$Pages, sapply(book_info_json$Authors, length)),
#   Release_Year = rep(book_info_json$`Release Year`, sapply(book_info_json$Authors, length)),
#   Price = rep(book_info_json$Price, sapply(book_info_json$Authors, length)),
#   ISBN = rep(book_info_json$ISBN, sapply(book_info_json$Authors, length))
# )
# # View the data frame
# print(book_df_json)

```

File links:-
<a href="https://raw.githubusercontent.com/topkelama/lfsStorage/main/info_books.xml">XML File</a>
<a href="https://raw.githubusercontent.com/topkelama/lfsStorage/main/info_books.json">JSON File</a>
<a href="https://raw.githubusercontent.com/topkelama/lfsStorage/main/info_books.html">HTML File</a>




Conclusion:- The JSON file did not have issue with the nested lists such as Author column, where there are multiple authors for a single Title. It automatically, did put in long format. However, with the xml and json, it did put multiple authors in a single row under the Authors column. Three dataframes have similarities but not identical. 


