---
title: "Data Science in Context"
author: "Chhiring Lama"
date: "2024-04-03"
output: ioslides_presentation
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r}
library(dplyr)
library(ggplot2)
library(jpeg)
library(tidyr)
library(forcats)
library(randomForest)
library(rpart)
library(rpart.plot)
```


## Decision Tree

DATA 607, Spring 2024<br />
Chhiring Lama

```{r, echo=FALSE}
url <- "https://media.githubusercontent.com/media/topkelama/lfsStorage/main/tree_ML.jpg"

```

```{r, echo=FALSE}
# Download the image
#download.file(url, destfile = "tree_ML.jpg", mode = "wb")
```

```{r, echo=FALSE}
# # Load the downloaded image
# img <- readJPEG("tree_ML.jpg")
# 
# # Get the dimensions of the image
# img_width <- dim(img)[2]
# img_height <- dim(img)[1]
# 
# # Create a plot with appropriate dimensions
# plot(1, type = "n", xlab = "", ylab = "", xlim = c(0, img_width), ylim = c(0, img_height), axes = FALSE)
# 
# # Display the image
# rasterImage(img, 0, 0, img_width, img_height)
```

## What is a Decision Tree

- Imagine a tree where each branch represents a decision, and each leaf represents an outcome.
- In simple terms, decision trees help us understand which decisions to make based on certain conditions.


## How Does a Decision Tree Work?

- At each step, a decision tree looks at the data and asks questions to make the best choice.
- It splits the data into smaller groups based on features that help classify or predict outcomes.

## Splitting Criteria

- The tree decides which questions to ask first based on how much they help to separate the data into pure groups.
- It's like picking the most useful questions to ask to get the clearest answers.

## Load the data from github

```{r, echo=TRUE}
loan_url <- "https://media.githubusercontent.com/media/topkelama/lfsStorage/main/loan_eligibility.csv"
loana_df<- read.csv(loan_url)
```

## Set a seed of random nuber and shuffle the data

```{r, echo=TRUE}
set.seed(1234)  # Set a seed for reproducibility 
shuffled_loan <- loana_df[sample(nrow(loana_df)), ] 
```

## Data Transformation

```{r, echo=TRUE}
# Drop variables
clean_loan <- shuffled_loan %>%
  select(-c(Loan.ID, Customer.ID)) %>%
  # Convert to factor level
  mutate(Term = factor(Term),
         Years.in.current.job = factor(Years.in.current.job),
         Home.Ownership = factor(Home.Ownership),
         Purpose = factor(Purpose),
         Bankruptcies = factor(Bankruptcies),
         Tax.Liens = factor(Tax.Liens)) %>%
  na.omit()
```
## Check current loan status

```{r, echo=TRUE}
clean_loan <- clean_loan %>%
  mutate(Loan_Status = ifelse(Current.Loan.Amount < 500000, "Approved", "Denied"))
```

## Create Fucntion for Train and Test

```{r, echo=TRUE}
create_train_test <- function(data, size = 0.8, train = TRUE) {
    n_row <- nrow(data)  
    total_row <- round(size * n_row) 
    
    train_sample <- sample(n_row, total_row) 
    
    if (train == TRUE) {
        return(data[train_sample, ])  
    } else {
        test_sample <- sample(setdiff(1:n_row, train_sample), n_row - total_row)
        return(data[test_sample, ])  
    }
}
```

## Dataset Splitting into 80/20

```{r, echo=TRUE}
train_set <- create_train_test(clean_loan, size = 0.8, train = TRUE)
dim(train_set)

test_set <- create_train_test(clean_loan, size = 0.8, train = FALSE)
dim(test_set)
```

## Define the formula for the decision tree model

```{r, echo=TRUE}
formula <- Loan_Status ~ Credit.Score + Years.in.current.job + Home.Ownership
```

## Build the decision tree model using rpart

```{r, echo=TRUE}
decision_tree <- rpart(formula, data = train_set, method = "class")
```

## Print the summary of the decision tree model

```{r, echo=TRUE}
printcp(decision_tree)
```

## Slide with Plot

```{r, echo=TRUE}
# Plot the decision tree
rpart.plot(decision_tree, uniform = TRUE, main = "Decision Tree Model")


```

## Evaluate the model on the testing set <br />
 And Calculate accuracy

```{r,echo=TRUE}
predictions <- predict(decision_tree, test_set, type = "class")

accuracy <- mean(predictions == test_set$Loan_Status)
print(paste("Accuracy of the decision tree model:", round(accuracy * 100, 2), "%"))
```

## The End


## Citation:- 

Personal loan modeling. (2020, March 10). Kaggle. https://www.kaggle.com/datasets/teertha/personal-loan-modeling

Memberdev_Admin. (2023, March 31). Om Swami Official Writings & Courses. os.me - a Spiritual Home. https://os.me/short-stories/the-upside-down-tree/





