---
title: "Global Baseline Estimate"
author: "Chhiring Lama"
date: "2024-04-01"
output: html_document
---

## Implementing Global Baseline Estimate Recommender System in R

# Load Libraries

```{r echo=TRUE}
#library(tidyverse)
library(dplyr)
library(DBI)
library(RMySQL)
```

```{r, echo=FALSE}
Sys.setenv(PASSWORD_DB = "lama65")
```

## Connect to MySQL Database
 password has been masked

```{r echo=TRUE}
lamadb <- dbConnect(RMySQL::MySQL(), 
                 user = "chhiring.lama65", 
                 password = Sys.getenv("PASSWORD_DB"),
                 dbname = "chhiring.lama65",
                 host = "cunydata607sql.mysql.database.azure.com")
```

# List the tables
```{r echo=TRUE}
dbListTables(lamadb)
```

# Show Column Names using dbListFields function (assigned it to column_names)
```{r echo=TRUE}
table_name <- "movie_rating"  
column_names <- dbListFields(lamadb, table_name)
```

# Construct the SQL query using sprintf function and then retrieve the data from MySQL database using dbGetQuery function
```{r echo=TRUE}
query <- sprintf("SELECT * FROM %s", table_name)
table_data <- dbGetQuery(lamadb, query)

```


```{r echo=TRUE}
table_data
```

# Import the table from MySQL back to R

```{r echo=TRUE}
table_name <- "movie_rating"

query <- paste("SELECT * FROM", table_name)

table_data <- dbGetQuery(lamadb, query)
```

```{r echo=TRUE}
table_data
```
## Implement Global Baseline Estimate 
# Calculate global mean rating

```{r, echo=TRUE}
str(table_data)
```

 Convert list columns to numeric vectors
```{r, echo=TRUE}
table_data[, 5:9] <- lapply(table_data[, 5:9], function(x) unlist(x))
```


 Convert rating columns to numeric
```{r, echo=TRUE}

table_data[, 5:9] <- lapply(table_data[, 5:9], as.numeric)
```

# Calculate global mean rating
```{r, echo=TRUE}
global_mean <- mean(unlist(table_data[, 5:9]), na.rm = TRUE)
```

# Calculate mean rating per movie

```{r, echo=TRUE}

movie_means <- colMeans(table_data[, 5:9], na.rm = TRUE)
```

# Find the baseline rating
```{r, echo=TRUE}
# Subtract movie means from global mean
baseline_ratings <- global_mean - movie_means
```

# Fill NA values in each column with the corresponding baseline rating

```{r, echo=TRUE}

table_data$The_Creator[is.na(table_data$The_Creator)] <- baseline_ratings["The_Creator"]
table_data$The_Holdovers[is.na(table_data$The_Holdovers)] <- baseline_ratings["The_Holdovers"]
table_data$The_Marvels[is.na(table_data$The_Marvels)] <- baseline_ratings["The_Marvels"]
table_data$Wonka[is.na(table_data$Wonka)] <- baseline_ratings["Wonka"]
table_data$Society_of_the_Snow[is.na(table_data$Society_of_the_Snow)] <- baseline_ratings["Society_of_the_Snow"]

```

```{r, echo=TRUE}
baseline_ratings

```



```{recho=TRUE}
# Display updated table_data
print(table_data)

```
# Write the dataframe to MySQL and create a new table as 'gbe_updated'

```{r, echo=TRUE}

dbWriteTable(lamadb, "gbe_updated", table_data, overwrite = TRUE)
```
# Best security practice, disconnect from the database

```{r, echo=TRUE}
dbDisconnect(lamadb)
```

